\documentclass{beamer}

\begin{document}

\title{Bayesian Analysis of Data}
\subtitle{Key Steps, following Krushcke, 2015}
\author{DRME}
\date{\today}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}{Bayesian Analysis Steps}
  \begin{enumerate}
    \item Identify relevant data and measurement scales.
    \item Define a descriptive model for the data.
    \item Specify a prior distribution on parameters.
    \item Use Bayesian inference to allocate credibility.
    \item Check posterior predictions against data.
  \end{enumerate}
\end{frame}

\section{Step 1: Identify Relevant Data}

\begin{frame}{Step 1: Identify Relevant Data}
  \begin{itemize}
    \item Determine measurement scales.
    \item Identify predictor and response variables.
  \end{itemize}
\end{frame}

\section{Step 2: Define Descriptive Model}

\begin{frame}{Step 2: Define Descriptive Model}
  \begin{itemize}
    \item Choose a mathematical form for the model.
    \item Ensure parameters align with theoretical goals.
  \end{itemize}
\end{frame}

\section{Step 3: Specify Prior Distribution}

\begin{frame}{Step 3: Specify Prior Distribution}
  \begin{itemize}
    \item Obtain audience approval for the prior.
    \item Ensure the prior reflects prior knowledge.
  \end{itemize}
\end{frame}

\section{Step 4: Bayesian Inference}

\begin{frame}{Step 4: Bayesian Inference}
  \begin{itemize}
    \item Use data to update beliefs (Bayesian inference).
    \item Interpret the posterior distribution.
  \end{itemize}
\end{frame}

\section{Step 5: Posterior Predictive Check}

\begin{frame}{Step 5: Posterior Predictive Check}
  \begin{itemize}
    \item Verify that posterior predictions match observed data.
    \item Consider alternative descriptive models if needed.
  \end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}
  \begin{itemize}
    \item Recap key steps in Bayesian analysis.
    \item Emphasize the importance of model checking.
  \end{itemize}
\end{frame}

\end{document}

