\documentclass{article}
\usepackage{graphicx}

\title{Probability and Statistics: Outline}
\author{DRME}
\date{2024-01-20}

\begin{document}

\maketitle

\section*{Understanding the Basics}

\subsection*{Overview of the Concept: The Set of All Possible Events}
\begin{itemize}
  \item Fundamental concept in probability theory.
  \item Represents all potential outcomes in an experiment.
  \item Significance in defining sample space and event probabilities.
\end{itemize}

\subsection*{Examples: Coin Toss and Rolling a Die}
\begin{itemize}
  \item Coin Toss: \{Heads, Tails\}
  \item Die Roll: \{1, 2, 3, 4, 5, 6\}
\end{itemize}

\subsection*{Coin Flips: Why You Should Care}
\begin{itemize}
  \item Illustration using coin flips.
  \item Linking the concept to real-world scenarios.
\end{itemize}

\subsection*{Probability: Outside or Inside the Head}
\begin{itemize}
  \item Outside the head: Long-run relative frequency.
  \item Inside the head: Subjective belief.
  \item Probabilities assign numbers to possibilities.
\end{itemize}

\subsection*{Understanding Probabilities: Coin Toss, Die Roll, Card Drawing}
\begin{itemize}
  \item Coin Toss: 0.5 for both Heads and Tails.
  \item Die Roll: \( \frac{1}{6} \) for each number.
  \item Card Drawing: \( \frac{4}{52} \) for Aces, \( \frac{26}{52} \) for red cards.
\end{itemize}

\section*{Probability Distributions: Discrete and Continuous}
\begin{itemize}
  \item Overview of probability distributions.
  \item Discrete: PMF (Bernoulli, Binomial, Poisson).
  \item Continuous: PDF (Uniform, Exponential, Normal).
\end{itemize}

\subsection*{Properties of PDFs and Focus on Normal Distribution}
\begin{itemize}
  \item Emphasis on normalization and non-negativity.
  \item Special focus on the normal distribution.
\end{itemize}

\subsection*{Mean, Variance, and Highest Density Interval (HDI)}
\begin{itemize}
  \item Mean as minimized variance.
  \item Definition and importance of HDI.
\end{itemize}

\section*{Two-Way Distributions: Conditional Probability and Independence}
\begin{itemize}
  \item Conditional Probability: \( P(A | B) = \frac{P(A \cap B)}{P(B)} \)
  \item Independence: \( P(A \cap B) = P(A) \cdot P(B) \)
\end{itemize}

\section*{Joint Probability and Applications}
\begin{itemize}
  \item Definition and interpretation.
  \item Applications in finance, medicine, and weather.
\end{itemize}

\subsection*{Rules of Factoring Joint Probabilities}
\subsubsection*{1. Product Rule}
\[ P(A \cap B) = P(A|B) \cdot P(B) \]
\begin{itemize}
  \item Genetic inheritance and species coexistence examples.
\end{itemize}

\subsubsection*{2. Chain Rule}
\[ P(A \cap B \cap C) = P(A|B \cap C) \cdot P(B|C) \cdot P(C) \]
\begin{itemize}
  \item Application in gene expression and ecosystem dynamics.
\end{itemize}

\subsubsection*{3. Marginalization}
\[ P(A) = \sum_{B} P(A \cap B) \]
\begin{itemize}
  \item Example with a joint probability distribution table.
\end{itemize}

\end{document}
